\chapter{SFAC Language Reference}
\label{cha:function}
This chapter describes the functions available in the SFAC interface. In the
documentation of each function, the arguments in brackets are optional,
arguments separated by ``$\mid$'' are alternative forms of calling syntax,
``...'' in the argument list denotes variable number of arguments, and keyword
arguments are indicated by \var{key=arg} pair.

\section{Variables}
One can use variables instead of explicitly typed arguments of any SFAC
function. This can be convenient if the same argument is used in more than one
place. A variable name may contain any alphanumeric characters but must begin
with a letter. Function names listed in Sec.~\ref{sec:fac} cannot be used as
variables names for obvious reasons. Once assigned, a variable is accessed by
prepending the dollar sign (\$) to its name. For example,
\begin{verbatim}
a = 2
Print($a)
2
\end{verbatim}
A variable may be either a scalar or a list (one-dimensional array), e.g.,
\begin{verbatim}
vv = [1,2,3]
Print($vv)
[1,2,3]
\end{verbatim}

\section{Functions}
\label{sec:fac}

The module contains the following functions.

\begin{fundesc}{AITable}{fn, b, f\opt{, c}}
Calculate the autoionization rates between the bound configuration group
\var{b} and the free configuration group \var{f}. The results are saved in
file \var{fn}. The optional channel number \var{c} can be supplied as an
identification of the transition array.
\end{fundesc}

\begin{fundesc}{AITableMSub}{fn, b, f\opt{, c}}
Calculate the magnetic sublevel autoionization rates and dielectronic capture
strength between the bound configuration group \var{b} and the free
configuration group \var{f}. The results are saved in file \var{fn}. The
optional channel number \var{c} can be supplied as an identification of the
transition array.
\end{fundesc}

\begin{fundesc}{AppendTable}{fn}
By default, when a new script is executed, existing binary files are
overwritten. If instead the new data should be appended to the file, use 
this function to set the append flag.
\end{fundesc}

\begin{fundesc}{Asymmetry}{fn, s\opt{, m}}
Calculate the photoionization asymmetry parameters for given relativistic
subshells. \var{s} is a string which gives the subshells. It should be in
spectroscopic notation, e.g., '1s' for $1s_{1/2}$, '2p' for $2p_{1/2,3/2}$
shells, '3p-' for $3p_{1/2}$ and '3p+' for $3p_{3/2}$, etc. The optional
\var{m} specifies the maximum multipole expansion, counting in the order E1,
M1, E2, M2, ..., therefore, $m=1$ includes only E1, $m=2$ includes E1 and M1,
etc. Default is $m=1$. The results are stored in file \var{fn}. For each
subshell, the output starts with one line indicating which subshell it is,
its $nlj$ values, the ionization energy, the number of energy points, and the
value of \var{m} used. It is followed by a block, which contains one line for
each energy point and tabulates the electron energy, photon energy, total
photoionization cross section ($10^{-20}$ cm$^2$), cross section for electron
direction perpendicular to that of photon, the total radiative recombination
cross section, the cross section at 90$^\circ$, and the ratio of
$\sigma_\perp$/$\sigma_\parallel$, where $\sigma_\perp$ is the 90$^\circ$
radiative cross section for photons polarized in the direction perpendicular
to the electron direction, and $\sigma_\parallel$ is the 90$^\circ$ cross
section for photons polarized in the direction parallel to the electron
direction. The last piece of data for each subshell are the $B_\lambda$ and
$B_\lambda^\phi$ parameters defined as
\begin{equation}
\frac{d\sigma}{d\Omega}=\frac{\sigma}{4\pi}\left[\sum_{\lambda\ge 0} B_\lambda
  P_{\lambda}(\cos\theta) - \sum_{\lambda\ge 2} \lambda^{-1}(\lambda-1)^{-1}
  B_\lambda^\phi P_{\lambda}^2(\cos\theta)\cos(2\phi)\right],
\end{equation}
where $\sigma$ is the total cross section, $\theta$ is the angle between the
electron and photon directions, $\phi$ is the azimuth angle of polarization
direction of the photon relative to the electron direction.
\end{fundesc}

\begin{fundesc}{AvgConfig}{c}
Setup a mean configuration for the optimization of central potential as
specified by a string \var{c}. The format of the string is the same as in the
function \key{Config}, except that the occupation can be a non-integer number
in this routine. The mean configuration setup by this function is effective
only if the function \key{OptimizeRadial} is called with no
arguments. Otherwise, the configurations given in that function are used to
generate the mean configuration automatically. It is important that this
function be called before \key{OptimizeRadial} and after
\key{ConfigEnergy(0)}, if the later is used.
\end{fundesc}

\begin{fundesc}{BasisTable}{fn\opt{,m}}
Print out a table of basis wavefunctions and mixing coefficients in the file
\var{fn}. If \var{m} is 0, then the basis table for the ordinary atom is
given, otherwise, the basis table for atom in magnetic and electric fields are
given.
\end{fundesc}

\begin{fundesc}{CETable}{fn, low, up}
Calculate the collision strength for the excitation of states in the
configuration group list \var{low} to those in the group list\var{up}. The
results are saved in the file \var{fn}.
\end{fundesc}

\begin{fundesc}{CETableEB}{fn, low, up\opt{, m}}
Calculate the collision strength for the excitation of states in the
configuration group list \var{low} to those in the group list\var{up}, but for
atoms in magnetic and electric fields. The
results are saved in the file \var{fn}. If \var{m} is 0, then incident
electron is assumed to be isotropic, otherwise, cross sections at different
incident directions are calculated.
\end{fundesc}

\begin{fundesc}{CETableMSub}{fn, low, up}
Calculate the magnetic sublevel collision strength for the excitation of
states in the configuration groups \var{low} to those in the groups
\var{up}. The results are saved in the file \var{fn}.
\end{fundesc}

\begin{fundesc}{CITable}{fn, b, f}
Calculate the collision strength for the ionization of states in the bound
configuration groups \var{b} to those in the free configuration groups
\var{f}. The results are saved in the file \var{fn}.
\end{fundesc}

\begin{fundesc}{CITableMSub}{fn, b, f}
Calculate the magnetic sublevel collision strength for the ionization of
states in the bound configuration groups \var{b} to those in the free
configuration groups \var{f}. The results are saved in the file \var{fn}.
\end{fundesc}

\begin{fundesc}{CheckEndian}{\opt{fn}}
Check the byte order of database file \var{fn}. It returns 0 for little endian
and 1 for big endian. If the optional file name \var{fn} is omitted, the
endian for the current platform is returned.
\end{fundesc}

\begin{fundesc}{Closed}{s, ...}
Specify the closed shells in the electronic configurations. It takes variable
number of arguments, each of them is a non-relativistic or relativistic shell
in the spectroscopic notation. e.g., \key{2s} for $2s$ shell, \key{2p-} for
$2p_{1/2}$ shell, and \key{2p+} for $2p_{3/2}$ shell.
\end{fundesc}

\begin{fundesc}{Config}{c, ..., group=g $\mid$ g, c, ...}
Add one or more configurations to the configuration group \var{g}. In the
first form, the group name \var{g} is given as a keyword, while in the second
form, the first argument must be a group name instead of a configuration. It
takes one or more strings for the configuration specification. A configuration
\var{c} is a string comprised of one or more non-relativistic or relativistic
shells in spectroscopic notation separated by white spaces. e.g.,
\key{2[p+]3} is a $2p_{3/2}$ shell with 3 electrons. If a \key{*} is given
instead of the orbital angular momentum symbol, configurations with all
legitimate values are generated. It is also possible to use \key{[s,p,d]} to
indicate that the orbital angular momentum may take $s$, $p$, or $d$
values. For $l > 20$, no spectroscopic symbol is available, it is specified as
[$l$] such as [21,22] for $l$ = 21 and 22 shells. This numerical notation also
works for $l \le 20$ shells, for which spectroscopic symbols \key{s, p, d, f,
g, h, i, k, l, m, n, o, q, r, t, u, v, w, x, y, z} are available. The bracket
for the orbital angular momentum can be omitted if it comprises of a single
character. Otherwise, the bracket must be present. For example, $2p-2$ is
illegal, but $2[p-]2$ is. Each shell may be followed by multiple conditions on
the occupation number, separated by ``;'', e.g., $3*10;3s>0;3p>5$ generate
configurations that have at least 1 electron in the $3s$ and 6 electrons in
the $3p$ shells. The logical relations allowed in conditions include $=$, $>$,
and $<$. 
\end{fundesc}

\begin{fundesc}{ConfigEnergy}{m\opt{, n\opt{, g, ...}}}
This function should be called twice just before (with \var{m} = 0) and after
(with \var{m} = 1) \key{OptimizeRadial} if used. If \key{AvgConfig} is
called, then \key{ConfigEnergy} with \var{m}=0 must be called before that.
The call with \var{m} = 0
performs a radial optimization for the configuration groups given by the list
\var{g}, and calculate the average energy of each configuration under
such potentials. Multiple optimizations are performed if more than one list
are given. If none is given, the optimization are carried out for each
configuration group. If \var{m} = 0, one may also
specify an integer \var{n} to indicate that \key{RefineRadial(n)} should be
called after the optimization. The call with \var{m} = 1 does not accept
additional arguments. It recalculates the average energy of each configuration
under the potential obtained by \key{OptimizeRadial} issued by the user, The
diagonal elements of the Hamiltonian calculated in the \key{Structure} call is
then adjusted by the difference of the two average energies for each
configuration. The purpose of this routine is to remove some of the errors in
the level energies introduced by using a single central potential for all
configurations.
\end{fundesc}

\begin{fundesc}{CorrectEnergy}{fn, nmin $\mid$ ilev, e, nmin}
Correct the energies of certain levels by given amount. This is used if the
exact energy of some levels is critical. In the first form, the indexes and the
energy corrections are listed as two columns in the file \var{fn}. In the
second form, the indexes are given in the list \var{ilev}, and the
energy corrections are given in the list \var{e}. Only levels with valence
electron in $n\ge nmin$ are corrected, if these levels are not constructed
with \key{RecStates}. The corrections are given in
units of eV. For the first level in the correction list, the amount is added
to the energies calculated by \cFAC, and that level is taken as the reference
level for all other entries in the list. The correction energies in the list
\var{e} for these other entries with \var{ilev}$\ge 0$ are the desired energies
relative to the reference level, i.e., they replace the energies calculated by
\cFAC. However, if \var{ilev}$<0$, then these corrections are also interpreted
as shifts to be added to the energies calculated by \cFAC.
\end{fundesc}

\begin{fundesc}{CutMixing}{g0, g1\opt{, c}}
For each levels in the group list \var{g0}, eliminate all mixing components
that are not in the group list \var{g1} or the mixing coefficients less than
\var{c}.
\end{fundesc}

\begin{fundesc}{Exit}{}
Exit \cFAC.
\end{fundesc}

\begin{fundesc}{GetConfigNR}{c, ...}
This functions returns a list of non-relativistic configurations corresponding
to the supplied configuration strings, which may contain wild casts as in the
function \key{Config}. It is useful, e.g., when one wants to know the
non-relativistic configurations of \key{1*2 2*2 3*1}. 
\end{fundesc}

\begin{fundesc}{GetPotential}{fn}
Print the radial potential obtained by \key{OptimizeRadial} to the file
\var{fn}. The file starts with the parameters for the analytic fit the to
potential, $\lambda$ and $a$, in the formula
\begin{equation}
V_0(r) = -\frac{Z}{r} + \frac{N-1}{r}\left(1-\frac{\exp(-\lambda
r)}{1+ar}\right). 
\end{equation}
After that, the mean configuration used to generate the potential is printed
in 3 columns representing the principle quantum number, the relativistic
angular quantum number $\kappa$, and the fractional occupation number
respectively. Finally, the file gives 8 columns which are $i$, $r$, $Z(r)$,
$V(r)$, $V_d(r)$, $V_e(r)$, $V_e^\prime(r)$, and $U(r)$, where $i$ is the index
for the radial grid, $r$ is the radial grid, $Z(r)$ is the nuclear charge at
radius $r$ taking into account the nuclear charge distribution, $V(r)$ is the
optimal potential, $V_d(r)$ is the direct interaction part of the potential,
$V_e(r)$ is the exchange interaction part of the potential, $V_e^\prime(r)$ is
the Slater approximation of the exchange interaction, and $U(r)$ is the
Uehling potential which approximates the vacuum polarization effects.
\end{fundesc}

\begin{fundesc}{Info}{}
Print out the version information of \cFAC and contact information of the
author. 
\end{fundesc}

\begin{fundesc}{JoinTable}{fn1, fn2, fn}
Join two binary files \var{fn1} and \var{fn2} to produce a single file
\var{fn}. \var{fn1} and \var{fn2} must have been produced on the same
platform, have the same type and for the same element.
\end{fundesc}

\begin{fundesc}{ListConfig}{\opt{fn, g}}
Print the configurations in the list \var{g} to file \var{fn}. If \var{fn} is
not given or if it is ``-'', the results are written to the stdout. If \var{g}
is not given, then all configurations currently defined are printed.
\end{fundesc}

\begin{fundesc}{MemENTable}{fn}
Build an energy level table in the memory for the \texttt{DB\_EN} type file
\var{fn}. This function must be called before any calls to \key{PrintTable}
in verbose mode.
\end{fundesc}

\begin{fundesc}{OptimizeRadial}{\opt{g\opt{, w}}}
Obtain the optimal radial potential based on the mean configuration generated
by the configuration group list \var{g} and the weight \var{w}, or if they are
absent in the call, by the mean configuration specified by \key{AvgConfig}.
\var{g} and \var{w} must be equal length lists if both present. \var{g}
is a list of configuration groups, and \var{w} is a list of weights for each
group when generating the mean configuration. If only \var{g} is present, each
configuration group is given an equal weight.
\end{fundesc}

\begin{fundesc}{Pause}{}
Pause execution of \cFAC.
\end{fundesc}

\begin{fundesc}{PrepAngular}{p\opt{, q}}
Precalculate the angular coefficients between states in \var{p} and
\var{q}. \var{p} and \var{q} are lists of configuration groups. If \var{q} is
not present, the angular coefficients are calculated between states in the
\var{p} list. Only $Z^L(\alpha,\beta)$ and $\tilde{a_\alpha}$ coefficients are
calculated. If the Bra and Ket states have the same number of electrons, $Z^L$
is calculated, otherwise $\tilde{a_\alpha}$ is calculated. This routine should
primarily be used when atomic states are constructed with \key{RecStates},
where the angular coefficients between the base states are used many times. It
is therefore more efficient to precalculate these coefficients.
\end{fundesc}

\begin{fundesc}{Print}{args}
Print out the string representation of \var{args}.
\end{fundesc}

\begin{fundesc}{PrintTable}{fnb, fna\opt{, v}}
Convert the binary database file \var{fnb} to the ASCII file \var{fna}. The
optional argument \var{v} = 1 requires the conversion be done in verbose
mode, otherwise it is done in simple mode. Note that before conversion in
verbose mode is carried out, one must call \key{MemENTable} first.
\end{fundesc}

\begin{fundesc}{RecStates}{fn, b, n}
Construct recombined states by adding a spectator electron with the principle
quantum number \var{n} onto the basis states in the configuration groups
\var{b}. The orbital angular momentum of the spectator electron is set by two
functions \key{SetRecPWLimits} and \key{SetRecSpectator}. The resulting energy
levels are saved in file \var{fn}.
\end{fundesc}

\begin{fundesc}{RefineRadial}{\opt{n\opt{, m}}}
This function may be called after \key{OptimizeRadial}, which performs a
minimization of the total energy of the mean configuration by adjusting the
parameters in the analytic central potential. \var{n} is the number of energy
evaluations allowed in the minimization, and \var{m} controls the print out
during the calculation. Default: \var{n} = 100, \var{m} = 0 (no print out).
\end{fundesc}

\begin{fundesc}{RRMultipole}{fn, b, f\opt{, m}}
Calculate the bound-free multipole matrix elements. Because \key{RRTable}
does not calculate the matrix elements directly, this function exists to
provide such data. The output are written to the file \var{fn} in ASCII
format. The format is described in the header of the file itself. The
remaining arguments are identical to those in \key{RRTable}
\end{fundesc}

\begin{fundesc}{RRTable}{fn, b, f\opt{, m}}
Calculate the bound-free differential oscillator strengths between the bound
configuration groups \var{b} and free groups \var{f}, which are related to
radiative recombination and photoionization cross sections. The optional
multipole type \var{m} is set to -1 (E1) by default. In almost all cases, no
other multipole types should be important. The results are saved in file
\var{fn}. 
\end{fundesc}

\begin{fundesc}{SetAICut}{c}
Set the autoionization rate cutoff threshold in the output. Only
autoionization rates greater than \var{c} a.u. are output. The default is
$10^{-16}$ a.u. or $\sim 4.13$ s$^{-1}$, if this routine is not called. 
\end{fundesc}

\begin{fundesc}{SetAngZCut}{c}
Set the cutoff threshold for the mixing basis in the calculation of recoupling
coefficients. Only the basis functions with mixing coefficients $>$\var{c} are
included. The default is $10^{-6}$ if this routine is not called.
\end{fundesc}

\begin{fundesc}{SetAtom}{asym\opt{, z\opt{, m\opt{,r}}}}
This function set the atomic element to \var{asym}, where \var{asym} is the
standard elemental symbol. The nuclear charge \var{z}, atomic mass \var{m},
and the nucleus radius of the element can be set optionally. If they are not
set, the standard values are used.
\end{fundesc}

\begin{fundesc}{SetBreit}{n}
Set the maximum principle quantum number of the orbitals for which the
Breit interaction should be included in the Hamiltonian. If \var{n} $<$ 0,
then the Breit interaction involving all bound and continuum states is
included. Default is 5.
\end{fundesc}

\begin{fundesc}{SetCEBorn}{e\opt{,x\opt{, x1}}}
\var{e} specifies the asymptotic energy where the plane-wave Born approximation is
expected to be valid, and is used to deduce the constant component of the
collision strength at high energies. If $e > 0$, it is in unit of the
characteristic transition energy of the array. If $e < 0$, then its absolute
value is the energy in unit of eV. $x$ and $x1$
set the energy boundary between the distorted-wave Born approximation and
plane-wave Born approximation for the collisional excitation. If $x > 0$ and
the scattered
electron energy is less than $xE_{b}$, DW method is chosen, where $E_{b}$ is
the binding energy of the electron being excited. 
If $x$ is negative, the switch between DW and PW occurs when the estimated
high partial-wave contributions becomes larger than $-x$ times the low
partial-waves calculated explicitly. If $x=0$, then PW is always used.
default for $x$ is $-0.5$, i.e., use DW when the estimated high partial-wave
contributions represents no more than 50\% of the explicitly calculated low
partial-wave sum. The contributions of high partial waves for monopole and
dipole transitions are calculated with coulomb-bethe approximation, their
cutoff values are treated differently with the switch \var{x1}, whose default
is $-1.0$.
\end{fundesc}

\begin{fundesc}{SetCEGrid}{g $\mid$ n\opt{, e0, e1}}
Set the collision energy grid for collisional excitation. In the first form, the
grid is given by a list \var{g}. In the second form, the grid is
constructed with \var{n} points, from \var{e0} to \var{e1}. The energies are
specified for the scattered electron energy, and in units of eV. This routine
does not need to be called. A default is constructed for a given transition
array with 6 points, minimum and maximum energies specified by
\key{SetCEGridLimits}. Calling this routine with \var{n}=0, reset the grid to
system default.
\end{fundesc}

\begin{fundesc}{SetCEGridLimits}{e0, e1}
Set the minimum and maximum collision energy for collisional excitation for
the automatic construction of the grid. They are in units of average threshold
energy of the transition array being considered. The default is 0.05 and 8.0
if this routine is not called.
\end{fundesc}

\begin{fundesc}{SetCELCB}{m}
Set the orbital angular momentum for Coulomb-Bethe approximation.
\end{fundesc}

\begin{fundesc}{SetCELMax}{m}
Set the maximum of orbital angular momentum for the partial-wave expansion in
collisional excitation.
\end{fundesc}

\begin{fundesc}{SetCELQR}{m}
Set the maximum orbital angular momentum for quasi-relativistic approximation
in collisional excitation. The default is 0, i.e., always use
quasi-relativistic approximation.
\end{fundesc}

\begin{fundesc}{SetCEQkMode}{mode}
Set the computation mode for the excitation radial integrals. \var{mode} may
be a string or an integer specifying the mode. These values are listed in the
variable \key{QKMODE}.
\end{fundesc}

\begin{fundesc}{SetCIEGrid}{g $\mid$ n\opt{, e0, e1}}
Set the collision energy grid for collisional ionization. In the first form, the
grid is given by a list \var{g}. In the second form, the grid is
constructed with \var{n} points, from \var{e0} to \var{e1}. The energies are
specified for the total energy of the final electrons, and in units of
eV. This routine does not need to be called. A default is constructed for a
given transition array with 6 points, minimum and maximum energies specified
by \key{SetCIEGridLimits}. 
Calling this routine with \var{n}=0, reset the grid to system default.
\end{fundesc}

\begin{fundesc}{SetCIEGridLimits}{e0, e1}
Set the minimum and maximum collision energy for collisional ionization for
the automatic construction of the grid. They are in units of average threshold
energy of the transition array being considered. The default is 0.05 and 8.0
if this routine is not called.
\end{fundesc}

\begin{fundesc}{SetCILCB}{m}
Set the orbital angular momentum for the Coulomb-Bethe approximation in DW
collisional ionization.
\end{fundesc}

\begin{fundesc}{SetCILevel}{m}
Set the level of configuration interaction space. By default, $m=0$, the
configuration space is determined by the configuration groups passed to the
\key{Structure} function. CI can be further refined by the value of $m$. If
$m=-1$, then no CI is included, the Hamiltonian is assumed to be diagonal. If
$m=1$, only CI within the same relativistic configuration is included. If
$m=2$, only CI within the same non-relativistic configuration is included. If
$m=3$, only CI within the same configuration group is included.
\end{fundesc}

\begin{fundesc}{SetCILMax}{m}
Set the maximum orbital angular momentum for the partial-wave expansion in DW
collisional ionization.
\end{fundesc}

\begin{fundesc}{SetCILMaxEject}{m}
Set the maximum orbital angular momentum for the ejected electron in DW
collisional ionization.
\end{fundesc}

\begin{fundesc}{SetCILQR}{m}
Set the orbital angular momentum for quasi-relativistic approximation in DW
collisional ionization.
\end{fundesc}

\begin{fundesc}{SetCITol}{m}
Set the tolerance factor in the partial-wave expansion of DW collisional
ionization. 
\end{fundesc}

\begin{fundesc}{SetCIQkMode}{mode}
Set the computation mode for the ionization radial integrals. \var{mode} may
be a string or an integer specifying the mode. These values are listed in the
variable \key{QKMODE}.
\end{fundesc}

\begin{fundesc}{SetFields}{b, e, a\opt{, m}}
Set the magnetic and electric fields. \var{b} is the magnetic fields in Gauss,
\var{e} is the electric fields in Volts/cm. \var{a} is the angle between the
magnetic and electric fields. If the optional \var{m} is 1, then the
diamagnetic effects is ignored in the Hamiltonian.
\end{fundesc}

\begin{fundesc}{SetHydrogenicNL}{\opt{n,\opt{l}}}
Set the principle quantum number \var{n} and the orbital angular momentum
\var{l}, beyond which, the hydrogenic approximation for the E1 multipole
integrals should be used. If this routine is not called or the argument is not
given, the default is \var{n}=8, and \var{l}=7. 
\end{fundesc}

\begin{fundesc}{SetIEGrid}{g $\mid$ n\opt{, e0, e1}}
Set the ionization threshold energy grid for the collisional ionization. In
the first form, the grid is given by a list \var{g}. In the second
form, the grid is constructed with \var{n} points from \var{e0} to
\var{e1}. This routine does not need to be called. A 3 point grid is
constructed according to the transition array being considered by default.
Calling this routine with \var{n}=0, reset the grid to system default.
\end{fundesc}

\begin{fundesc}{SetMaxRank}{k}
Set the maximum rank in the expansion of Slater integrals. The default is 6,
if this routine is not called.
\end{fundesc}

\begin{fundesc}{SetMixCut}{c}
Set cutoff threshold of the mixing basis in the wavefunction. Only the basis
with mixing coefficients greater than \var{c} are included in the wavefunction
expansion. Default is $10^{-5}$.
\end{fundesc}

\begin{fundesc}{SetMS}{nms, sms}
Set flags for normal mass shift and specific mass shift contributions to the
Hamiltonian. 0---disable, 1---enable. Defaults: 1.
\end{fundesc}

\begin{fundesc}{SetOptimizeControl}{t, s, m\opt{, p}}
Set the options for radial potential optimization. \var{t} is the tolerance
for the self-consistent field iteration. \var{s} is the stabilizer for the
iteration, a number from 0 to 1. \var{m} is the maximum number of iterations
allowed. \var{p} 
specifies whether diagnostic information should be printed out during the
optimization. This routine does not need to be called. The default for \var{t}
is $10^{-6}$, \var{s} is determined dynamically according to the
type of ion, \var{m} is 100, and \var{p} is 0 for no printing out of
information. 
\end{fundesc}

\begin{fundesc}{SetOptimizeMaxIter}{m}
Set the maximum itneration in \key{OptimizeRadial}.
\end{fundesc}

\begin{fundesc}{SetOptimizePrint}{m}
Set the printing option in \key{OptimizeRadial}.
\end{fundesc}

\begin{fundesc}{SetOptimizeStabilizer}{m}
Set the stabilizer factor in \key{OptimizeRadial}.
\end{fundesc}

\begin{fundesc}{SetOptimizeTolerance}{m}
Set the tolerance factor in \key{OptimizeRadial}.
\end{fundesc}

\begin{fundesc}{SetPEGrid}{g $\mid$ n\opt{, e0, e1}}
Set the free electron energy grid for photoionization, radiative recombination
and autoionization. In the first form, the grid is given by a list
\var{g}. In the second form, the grid is constructed with \var{n} points from
\var{e0} to \var{e1}. The energies are in units of eV. This function does not
need to be called. A 6 point grid is constructed according to the transition
array being considered by default. Calling this function with \var{n}=0 reset
the grid to system default. 
\end{fundesc}

\begin{fundesc}{SetPEGridLimits}{e0, e1}
Set the minimum and maximum collision energy for photoionization and radiative
recombination for the automatic construction of the grid. They are in units of
average threshold energy of the transition array being considered. The default
is 0.05 and 8.0 if this routine is not called.
\end{fundesc}

\begin{fundesc}{SetRadialGrid}{n\opt{, r0\opt{,r1}\opt{,rmin}}}
Set the radial grid properties. \var{n} is the number of radial grid
points. It must be an even number and less than the macro
\key{MAXRP}. \var{r0} specifies the ratio of successive radial points near 
origin, which is approximately logarithmic. \var{r1} specifies the number of
mesh-points per oscillation wavelength for very high-$n$ orbitals at large 
radii. \var{rmin} divided by the nuclear charge is the starting point of the
radial mesh.
\end{fundesc}

\begin{fundesc}{SetRRTEGrid}{g $\mid$ n\opt{, e0, e1}}
Set the transition energy grid for photoionization and radiative
recombination. In the first form, the grid is given by a list
\var{g}. In the second form, the grid is constructed with \var{n} points from
\var{e0} to \var{e1}. This routine does not need to be called. For E1 type
transitions, the transition energy does not enter the calculation, a 1-point
grid is constructed by default. for other types of multipoles, a 3-point grid
is constructed. Calling this function with \var{n}=0 reset the grid to system default.
\end{fundesc}

\begin{fundesc}{SetRecPWLimits}{l0, l1}
Set the orbital angular momentum range for the spectator electron in the
recombined states to [\var{l0}, \var{l1}] inclusive. The default is [0, 12].
\end{fundesc}

\begin{fundesc}{SetRecPWOptions}{lmax}
Set maximum orbital angular momentum of the spectator electron in the
recombined states to \var{lmax}. The allowed values are also limited by the
setting of \key{SetRecPWLimits}. The default is 12.
\end{fundesc}

\begin{fundesc}{SetRecQkMode}{mode}
Set the computation mode for the photoionization and radiative recombination
radial integrals. \var{mode} may be a string or an integer specifying the
mode. These values are listed in the variable \key{QKMODE}.
\end{fundesc}

\begin{fundesc}{SetRecSpectator}{nmin\opt{, nfrozen}}
Set the minimum principle quantum numbers \var{nmin} and \var{nfrozen} for the
spectator electron. States with $n > nmin$ are constructed with
\key{RecStates} function, and those with $n > nfrozen$ are treated with frozen
core approximation. Default for both \var{nmin} and \var{nfrozen} are 8.
\end{fundesc}

\begin{fundesc}{SetScreening}{ns\opt{,c\opt{,k}}}
Set the orbital parameters of screening electrons. \var{ns} is a list of
integers which are the principle quantum numbers of the screening
orbitals. The optional \var{c} is the total charge to be screened, whose
default is 1.0. \var{k} is the either 1, $-1$, or 0. If \var{k}=$-1$, then
the \var{l=0} orbitals are used. If \var{k}=0, then the the \var{l}=\var{ns}/2
orbitals are used. If \var{k}=1, then the \var{l}=\var{ns}-1 nodeless orbital
is used. The default is \var{k}=1. This function is usually used when
additional screening charge is desired for the mean configuration generating
the optimal central potential. It is quite experimental, and therefore not
recommended for general use.
\end{fundesc}

\begin{fundesc}{SetSE}{n}
Set the maximum principle quantum number of the orbitals for which the
self-energy correction should be included in the Hamiltonian. Default is 5.
\end{fundesc}

\begin{fundesc}{SetSlaterCut}{k1, k2}
Set the calculation modes of Slater integrals. \var{k1} and \var{k2} are
orbital angular momentum values. When one of the orbitals has $l > k1$, then
exchange integrals are not calculated. When one has $l > k2$, the direct
integrals are evaluated with the multipole moments.
\end{fundesc}

\begin{fundesc}{SetTEGrid}{g $\mid$ n\opt{, e0, e1}}
Set the transition energy grid for collisional excitation. In the first form,
the grid is given by a list \var{g}. In the second form, the grid is
constructed with \var{n} points from 
\var{e0} to \var{e1}. This routine does not need to be called. A 3-point grid
is constructed by default. Calling this function with \var{n}=0 reset the grid
to system default. 
\end{fundesc}

\begin{fundesc}{SetTransitionGauge}{m}
Set the gauge for radiative transition.
\end{fundesc}

\begin{fundesc}{SetTransitionMaxE}{m}
Set the maximum rank of electric multipole transitions.
\end{fundesc}

\begin{fundesc}{SetTransitionMaxM}{m}
Set the maximum rank of magnetic multipole transitions.
\end{fundesc}

\begin{fundesc}{SetTransitionMode}{m}
Set the mode for the radiative transition.
\end{fundesc}

\begin{fundesc}{SetTransitionOptions}{g, m}
Set the options for the radiative transition calculation. \var{g} is the gauge
to be used, 1 for Coulomb gauge (velocity form) and 2 for Babushkin gauge
(length form, which is the default). \var{m} is the mode for the multipole
integral, 0 for fully relativistic and 1 for non-relativistic approximation
(the default is 1).
\end{fundesc}

\begin{fundesc}{SetUsrCEGrid}{g $\mid$ n\opt{, e0, e1}}
Set the user collision energy grid for collisional excitation. The collision
strengths on this grid are output. It is forced to be same as that set by
\key{SetCEGrid} if the \key{QKMODE} is \key{'exact'}.
\end{fundesc}

\begin{fundesc}{SetUsrCIEGrid}{g $\mid$ n\opt{, e0, e1}}
Set the user collision energy grid for collisional ionization. The collision
strengths on this grid are output. 
\end{fundesc}

\begin{fundesc}{SetUsrPEGrid}{g $\mid$ n\opt{, e0, e1}}
Set the user electron energy grid for photoionization and radiative
recombination. The bound-free differential oscillator
strengths on this grid are output. It is forced to be same as that set by
\key{SetPEGrid} if the \key{QKMODE} is \key{'exact'}.
\end{fundesc}

\begin{fundesc}{SetUTA}{m\opt{, ci}}
Set the flag for configuration average models. If $m=1$, all calculations are
carried out in the configuration average approximation. The radiative
transition rates output contains three additional fields, the transition
energy including the UTA shift,
the Gaussian standard deviation, and the correction to the line strengths due to
the configuration interaction within the same non-relativistic
configurations. If $m=0$, which is the default,
the usual detailed term accounting method is used. This function should be
called in the beginning of the script, before \key{Config} function, since it
disables the angular momentum coupling performed by \key{Config}. The optional
argument \var{ci} indicates whether the configuration interaction correction
factors should be included when reading the \key{DB\_TR} files. These 
correction factors are always calculated in \key{TransitionTable} irrespective
of the value of \var{ci}.
\end{fundesc}

\begin{fundesc}{SetVP}{vp}
Set the flag for vacuum polarization correction to the
Hamiltonian. 0---disable, 1---only include 2nd order term, 2---include the 4th
order term as well. Default is 2.
\end{fundesc}

\begin{fundesc}{SlaterCoeff}{fn, g, a, b}
Calculate the expansion coefficients of the exchange radial integral
in the Coulomb energy of each state in the
configuration list 
\var{g}. The Coulomb energy between electrons of a state can be expanded as
\begin{equation}
E = \sum g_kG^k(\alpha\alpha^\prime,\beta\beta^\prime) + \mbox{direct terms},
\end{equation}
where $\alpha$ and $\alpha^\prime$ are the interacting orbitals in the bra and
ket states, which have the same $l$ values. Because we work in the
$jj$-coupling basis, $\alpha$ and $\alpha^\prime$ may have different $j$
values. The results are stored in the text file \var{fn}. These coefficients
are useful in determining the radiative transition rates from level to average
configurations. \var{a} and \var{b} are the orbital lists contain $\alpha$ and
$\beta$, respectively. They are specified as configuration strings.
The format of the file is as follows. For each state, a line
starts with ``\#'' gives the level index, parity, $2J$-value, and the
configuration label. It is followed by a block of lines. In this block, the first
column is the level index. The 2rd, 3th, and 4th columns are the $n$, $kappa$,
$l$ values of the $\alpha$ orbital. The 5th column is either 0 or 1,
indicating whether $\alpha$ and $\alpha^\prime$ have the same $j$ values. The
6-9th  columns are the corresponding values for the $\beta$ orbital. The 10th
column is the expansion coefficients $g_1$. The 11th column is $g_2$. The 12th
column is the the number to be added to $g_1$ to form the relative intensities
of level to configuration transitions for dipole transitions. The 13th column
is the number to be added to $g_2$ to form relative intensities for quadrupole
transitions.
\end{fundesc}

\begin{fundesc}{StoreClose}{}
Close the database that has previously been initialized with \key{StoreInit}.
\end{fundesc}

\begin{fundesc}{StoreInit}{fn \opt{, reset}}
Initialize a new session in the SQLite database file \var{fn}. If the file does
not exist, it will be created. If the optional flag \var{reset} = 1 is present
and the database already exists, all existing sessions will be dropped.
\end{fundesc}

\begin{fundesc}{StoreTable}{fn}
Store data contained in the binary file \var{fn} into a database that has
previously been initialized with \key{StoreInit}.
\end{fundesc}

\begin{fundesc}{Structure}{fn, g\opt{, p\opt{, ip}}}
Diagonalize the Hamiltonian for configurations in the groups \var{g}. The
configurations in the optional groups \var{p} are allowed to interact with
\var{g} but only states within \var{g} are added to the energy level table. If
\var{ip}=0, the interaction between \var{g} and \var{p} are treated exactly,
if \var{ip}=1, this interaction is treated approximately in a way that the
non-diagonal elements within \var{p} are neglected. The energy levels are
output to the file \var{fn}.
\end{fundesc}

\begin{fundesc}{Structure}{p, j}
This is the second form of the \key{Structure} function. It distinguishes
itself from the first form because the first argument is an integer. It sets which
$\pi J$ symmetry and/or which level to include in the structure calculation. By
default, all symmetries are processed. But if this function is called with
$p=0$, or 1; $j \ge 0$ or $j$ is a list of integers, then only the specified
symmetry is processed. $j$ or integers in the list if $j$ is a list, is 
twice the actual value of the total angular momentum. The symmetry
restrictions specified here also applies to the \key{StructureMBPT} function.
\end{fundesc}

\begin{fundesc}{StructureEB}{fn, g}
Calculate the atomic structure for atom in magnetic and electric fields. The
levels belong to the configuration group \var{g} are allowed to mix in the
external fields.
\end{fundesc}

\begin{fundesc}{TransitionTable}{fn, low, up\opt{, m}}
Calculate the weighted oscillator strength and radiative transition rates from
states in \var{up} groups to states in \var{low} groups with multipole type
\var{m}. The default for \var{m} is -1, i.e., E1 transitions. The results are
saved to file \var{fn}.
\end{fundesc}

\begin{fundesc}{TRTable}{fn, low, up\opt{, m}}
Same as \key{TransitionTable}
\end{fundesc}

\begin{fundesc}{TRTableEB}{fn, lo, up\opt{, m}}
Calculate radiative transition rates between levels in external magnetic and
electric fields.
\end{fundesc}

\begin{fundesc}{WaveFuncTable}{fn, n, k\opt{, e}}
Print the radial wavefunction of the orbital with the principle quantum number
\var{n}, relativistic angular quantum number $\kappa$=\var{k} to the file
\var{fn}. If \var{n}=0, the orbital is a continuum state. In this case, the
optional \var{e} must be a positive number for the energy of the continuum
orbital in unit of eV.
\end{fundesc}
